\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage[round]{natbib}
\usepackage{hyperref}
\hypersetup{hidelinks}
\usepackage{microtype}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\ind}[1]{\mathbf 1\{#1\}}

\title{A Contragredient View of Conformal Placement\\and Steinitz Balancing}
\author{Peter Cotton\\\texttt{peter.cotton@microprediction.com}}
\date{July 2026}

\begin{document}
\maketitle

\begin{center}
\fbox{\parbox{0.8\textwidth}{\centering\textbf{ROUGH DRAFT --- NOT FOR DISSEMINATION}}}
\end{center}
\vspace{6pt}

\begin{abstract}
We give the validity of split conformal prediction an exact contragredient form: the placement
acceptance map is equivariant for the permutation action, and the Reynolds identity
$\langle P^{\vee}f,v\rangle=\langle f,Pv\rangle$ reads its orbit average either as moving the
evaluation functional or as averaging the induced acceptance vector over placements. A natural
primal discrepancy construction under the same action is Steinitz balancing: ordering a
zero-sum population so that every prefix sum stays bounded. Conformal placements form such a
population, and an explicit ordering keeps running coverage within $1/(2t)$ of the orbit level
$k/N$, exact once all $N$ candidate test placements have been traversed; the vector-valued
lemma balances several coverage levels along
one traversal. On the function--measure pairing, the measure-side counterpart
of calibration is balanced sampling and exact cubature. A transitive distributional symmetry,
exchangeability in the standard setting, turns the deterministic orbit average into the
coverage probability of the designated test placement.
\end{abstract}

\section{Introduction}\label{sec:intro}

Split conformal prediction validates its coverage claim by a symmetry argument over sample
slots. This note gives that argument an exact dual form and identifies the deterministic
construction that lives beside it under the same group action. The identity behind conformal
validity is a contragredient statement for the permutation action on slots
(Section~\ref{sec:contra}). A natural primal discrepancy construction under that action is
\emph{Steinitz balancing}: ordering a finite zero-sum population so that every prefix sum stays
uniformly bounded \citep{steinitz1913}. Conformal placements are themselves a scalar Steinitz
population (Section~\ref{sec:balanced}), and admit a sharper balanced ordering than the general
lemma provides. On the function--measure pairing, the measure-side counterpart of calibration
is balanced sampling and exact cubature (Section~\ref{sec:cubature}). Section~\ref{sec:bp}
records how the same two-ingredient structure appears in the dependent-data floor of
\citet{barberpananjady2026}, and Section~\ref{sec:rates} separates three rates that the shared
algebra can blur. The group-symmetry view of predictive inference is itself established:
\citet{dobriban2025symmpi} build predictive inference on general distributional group
symmetries, of which exchangeability is the permutation case. What we believe is new here is
narrower: the identification of the placement acceptance vector as a Steinitz population, and
the balanced traversals it admits, at one level and at several simultaneously. Herding \citep{welling2009}, whose $O(1/n)$ bookkeeping
resembles the constructions here, can be viewed as a with-replacement relaxation of the
finite balanced-permutation problem \citep{harvey2014}; it is confined to
Remark~\ref{rmk:herding} and a kernel question in Section~\ref{sec:cubature}.

\section{Conformal calibration and placement}\label{sec:conformal}

Fix $\alpha\in(0,1)$, write $p=1-\alpha$, let $S_1,\dots,S_n$ be calibration scores and
$S_{n+1}$ a test score, almost surely distinct, and set $k=\lceil p\,(n+1)\rceil$ and $N=n+1$.
Assume throughout that $\alpha\ge 1/N$, equivalently $k\le n$, so that the $k$-th order
statistic of the calibration scores exists; for $\alpha<1/N$ the usual convention
$q_n=+\infty$ applies and coverage is trivial. Split conformal declares coverage when
$S_{n+1}\le q_n$, with $q_n=S_{(k)}$ the $k$-th smallest calibration score. The scores are
taken exchangeable, as they are in split conformal conditionally on an independent proper
training sample once the fitted score rule is held fixed.

\begin{lemma}[Steering identity]\label{lem:steer}
With $q_n=S_{(k)}$, $h_q(s)=\ind{s\le q}$, and the scores distinct,
\[
\sum_{i=1}^{n}\bigl(h_{q_n}(S_i)-p\bigr)\;=\;k-pn\;\in\;[\,p,\;1+p\,).
\]
\end{lemma}

\begin{proof}
Exactly $k$ calibration scores satisfy $S_i\le S_{(k)}$; and $p(n+1)\le k<p(n+1)+1$.
\end{proof}

The material of this section is standard \citep{papadopoulos2002,vovk2005,lei2018}.

\begin{proposition}[Counting identity]\label{prop:count}
Let $s_1,\dots,s_{N}$ be distinct reals and, for $i\in\{1,\dots,N\}$, call placement $i$
\emph{covered} if $s_i$ is at most the $k$-th smallest of $\{s_j\}_{j\ne i}$, writing
$\chi_i(s)=\ind{s_i\le Q_k(s_{-i})}$ for the indicator. Then $\chi_i(s)=1$ iff the rank of $s_i$
among all $N$ values is at most $k$, and consequently $\sum_{i=1}^{N}\chi_i(s)=k$.
\end{proposition}

\begin{proof}
Placement $i$ is covered iff at most $k-1$ of the other values are smaller than $s_i$, iff its
overall rank is at most $k$; ranks are a permutation of $\{1,\dots,N\}$.
\end{proof}

\begin{corollary}[Marginal validity]\label{cor:valid}
If $(S_1,\dots,S_{N})$ is exchangeable with almost surely distinct components, then
$\Prob\{S_{N}\le q_n\}=k/N\ge 1-\alpha$.
\end{corollary}

\begin{proof}
The coverage event is the event that placement $N$ is covered. By exchangeability every
placement indicator has the same expectation; average Proposition~\ref{prop:count}.
\end{proof}

The control-theoretic reading of the calibration step is established in the online literature:
the level update of adaptive conformal inference \citep{gibbscandes2021} is integral feedback on
realized miscoverage, and \citet{angelopoulos2023pid} name the identification. There the
controlled quantity is a time average along a sequence; Lemma~\ref{lem:steer} is the batch
counterpart, a one-shot quantized correction: the rank count is exactly $k$, while the level
error $k/n-p$ is quantized at $O(1/n)$.

\section{The contragredient identity}\label{sec:contra}

Let $V=\R^{N}$, with $\mathfrak S_{N}$ acting by permuting slots,
$(\rho(\pi)v)_j=v_{\pi^{-1}(j)}$, and let $\rho^{\vee}(\pi)f=f\circ\rho(\pi^{-1})$ be the
contragredient action on the dual space $V^{*}$ (for the dual representation and the averaging
operator see, e.g., \citealp{serre1977}). The coordinate functionals $e^{*}_{i}$ and the
prefix-sum functionals $L_t=\sum_{j\le t}e^{*}_{j}$ are elements of $V^{*}$, and one has,
exactly,
\begin{equation}\label{eq:contra}
L_t\bigl(\rho(\pi)v\bigr)\;=\;\bigl(\rho^{\vee}(\pi^{-1})L_t\bigr)(v):
\end{equation}
permute the points holding the prefix functional fixed, or hold the points fixed and permute the
functional. Averaging over the group, let $P=\frac1{N!}\sum_{\pi}\rho(\pi)$ be the Reynolds
projector and $P^{\vee}=\frac1{N!}\sum_{\pi}\rho^{\vee}(\pi)$ its dual, so that
\begin{equation}\label{eq:reynolds}
\langle P^{\vee}f,\,v\rangle\;=\;\langle f,\,Pv\rangle
\qquad\text{for all } f\in V^{*},\ v\in V .
\end{equation}
For the permutation representation, $Pv=\bar v\,\mathbf 1$ and
$P^{\vee}e^{*}_{N}=\frac1N\mathbf 1^{*}$.

The object the projector acts on is not the score vector but the acceptance vector. Write
$\R^{N}_{\ne}$ for the score vectors with distinct entries and
$\chi:\R^{N}_{\ne}\to\{0,1\}^{N}\subset V$ for the placement-acceptance map of
Proposition~\ref{prop:count}. Since $\chi_i(s)$ depends only on the rank of $s_i$ among the
entries of $s$, the map is equivariant:
\begin{equation}\label{eq:equivariant}
\chi\bigl(\rho(\pi)s\bigr)\;=\;\rho(\pi)\,\chi(s)
\qquad\text{for all }\pi\in\mathfrak S_{N},\ s\in\R^{N}_{\ne}.
\end{equation}
The distinction matters because $\chi$ is nonlinear and is defined here only on score vectors
with distinct entries. The averaged score vector $Ps=\bar s\,\mathbf 1$ leaves that domain and
erases all rank information. One must therefore average the induced acceptance vectors
$\chi(\rho(\pi)s)$, rather than apply $\chi$ after averaging the scores.

\begin{theorem}[Orbit--contragredient form of conformal validity]\label{thm:orbit}
For every $s\in\R^{N}_{\ne}$,
\[
\frac1{N!}\sum_{\pi\in\mathfrak S_{N}}\chi_{N}\bigl(\rho(\pi)s\bigr)
\;=\;\langle e^{*}_{N},\,P\chi(s)\rangle
\;=\;\langle P^{\vee}e^{*}_{N},\,\chi(s)\rangle
\;=\;\frac1N\sum_{i=1}^{N}\chi_i(s)
\;=\;\frac{k}{N}.
\]
Consequently, whenever the law of $S$ is invariant under a subgroup of $\mathfrak S_{N}$ acting
transitively on the slots, in particular under exchangeability, the designated test placement
has coverage probability exactly $k/N$.
\end{theorem}

\begin{proof}
The first equality is \eqref{eq:equivariant} and linearity of $e^{*}_{N}$; the second is
\eqref{eq:reynolds}; the third is $P^{\vee}e^{*}_{N}=\frac1N\mathbf 1^{*}$; the fourth is
Proposition~\ref{prop:count}. For a transitive subgroup $H\le\mathfrak S_N$, set $P_H=\frac1{|H|}\sum_{h\in H}\rho(h)$;
transitivity gives $P_Hv=\bar v\,\mathbf 1$, so the same Reynolds calculation runs verbatim
with $P_H$ in place of $P$ and pins each $\E\chi_i(S)$ to $k/N$.
\end{proof}

Define also the centered placement residual vector $r(s)=(\chi_i(s)-k/N)_{i\le N}$;
Proposition~\ref{prop:count} says $\langle\mathbf 1^{*},r(s)\rangle=0$, i.e.
\begin{equation}\label{eq:orbitzero}
\langle P^{\vee}e^{*}_{N},\,r(s)\rangle\;=\;0
\qquad\text{for every distinct } s .
\end{equation}
The identity is exact and requires no analogy; it is the defining identity of the dual
representation, applied to the acceptance vector. It does not remove the distributional
assumption: Theorem~\ref{thm:orbit} is a statement about the whole orbit, and a transitive
symmetry of the law, exchangeability being the standard case, is what makes the designated
placement inherit the orbit average. The systematic development of predictive inference from
distributional group symmetries is due to \citet{dobriban2025symmpi}.

\section{Steinitz balancing}\label{sec:steinitz}

A natural primal discrepancy construction under the same $\mathfrak S_{N}$ action is the
ordering of a finite zero-sum population. Let $E$ be a $d$-dimensional normed vector space;
finite dimensionality is essential, as no constant independent of $N$ is available in general
otherwise. If $z_1,\dots,z_{N}\in E$ satisfy $\sum_i z_i=0$ and $\|z_i\|\le 1$, the Steinitz
lemma \citep{steinitz1913} supplies a permutation $\pi$ with every prefix bounded,
$\|\sum_{j\le t}z_{\pi(j)}\|\le C_E$ for all $t$, with $C_E$ depending on the space and norm
but not on $N$ (a bound linear in $d$ holds for arbitrary norms; \citealp{grinberg1980}); hence
running averages within $C_E/t$ of zero. This is the bounded-ledger $O(1/t)$ phenomenon,
obtained by an element of the same permutation group that Section~\ref{sec:contra} acts with.

Two precisions. First, the vector-valued setting scalarizes cleanly: on $W=E^{N}$ define, for
each $\lambda\in E^{*}$, the functional $L_{t,\lambda}(z)=\lambda(\sum_{j\le t}z_j)\in W^{*}$;
then \eqref{eq:contra} holds verbatim for $L_{t,\lambda}$, and
$\|\sum_{j\le t}z_{\pi(j)}\|=\sup_{\|\lambda\|_{*}\le 1}L_{t,\lambda}(\rho(\pi^{-1})z)$
recovers the norm bound. Second, Steinitz balancing is not itself the contragredient of
conformal averaging. Every permutation, balanced or not, satisfies the transport identity
\eqref{eq:contra}; the Reynolds average gives the exact orbit identity of
Theorem~\ref{thm:orbit}; what Steinitz contributes is the choice of one $\pi$, data-dependent,
that makes the whole nested family of prefix functionals small at once.

For the relation to herding, see Remark~\ref{rmk:herding}.

\section{Balanced conformal placements}\label{sec:balanced}

Conformal placements are themselves a scalar Steinitz population: by
Proposition~\ref{prop:count} the residuals $r_i(s)=\chi_i(s)-k/N$ satisfy $\sum_i r_i(s)=0$ and
$|r_i(s)|\le 1$. Because they take only two values, the ordering can be made sharper than the
general lemma provides. The balancing step is a statement about binary vectors, so we state it
that way; the conformal content is that Proposition~\ref{prop:count} produces such a vector.

\begin{lemma}[Balanced ordering of a binary population]\label{lem:binary}
Let $b\in\{0,1\}^{N}$ with $\sum_i b_i=k$. There is a permutation $\pi\in\mathfrak S_{N}$ such
that, simultaneously for every $t=1,\dots,N$,
$\bigl|\sum_{j\le t}b_{\pi(j)}-tk/N\bigr|\le\tfrac12$.
\end{lemma}

\begin{proof}
Define the balanced binary word
$w_t=\lfloor tk/N+\tfrac12\rfloor-\lfloor (t-1)k/N+\tfrac12\rfloor$. Since $0\le k/N\le 1$,
each $w_t\in\{0,1\}$, and $\sum_{j\le t}w_j=\lfloor tk/N+\tfrac12\rfloor$, so
$|\sum_{j\le t}w_j-tk/N|\le\tfrac12$ for every $t$, with exactly $k$ ones among
$w_1,\dots,w_N$. Send the indices with $b_i=1$ to the positions with $w_t=1$ and the rest to
the rest.
\end{proof}

\begin{corollary}[Balanced placement ordering]\label{prop:balanced}
For every distinct score vector $s$ there is a permutation $\pi\in\mathfrak S_{N}$ such that,
simultaneously for every $t=1,\dots,N$,
\[
\Bigl|\frac1t\sum_{j=1}^{t}\chi_{\pi(j)}(s)-\frac{k}{N}\Bigr|\;\le\;\frac{1}{2t},
\]
with equality to zero at $t=N$, and hence
$\bigl|\frac1t\sum_{j\le t}\chi_{\pi(j)}(s)-p\bigr|<\frac1{2t}+\frac1N$ against the nominal
level $p$.
\end{corollary}

\begin{proof}
Apply Lemma~\ref{lem:binary} to $b=\chi(s)$, using Proposition~\ref{prop:count}; the second
bound adds $|k/N-p|<1/N$.
\end{proof}

The lemma is one-dimensional balanced rounding; prefix bounds of this type are classical, going
back to the chairman assignment problem \citep{tijdeman1980}, and no specifically conformal
structure enters once the count $\sum_i\chi_i(s)=k$ is known. The contribution is the
identification: the conformal placement vector is an object to which balanced-word machinery
applies, and the traversal it yields tracks the orbit level $k/N$, not the nominal $p$, exactly
once all $N$ candidate test placements have been traversed, and within $1/(2t)$ before that.
The traversal orders the $N$ coordinates $\chi_1(s),\dots,\chi_N(s)$, that is the $N$-element
coordinate orbit $\mathfrak S_N/\mathfrak S_{N-1}$, not the $N!$-element score orbit; the
group average of Theorem~\ref{thm:orbit} reduces to this $N$-placement average because each
candidate test coordinate occurs equally often. In the notation of Section~\ref{sec:contra}
the two readings of the traversal are literal:
$\langle L_t,\rho(\pi^{-1})r(s)\rangle=\langle\rho^{\vee}(\pi)L_t,r(s)\rangle$ --- move the
residual vector through the permutation holding the prefix test fixed, or hold the residuals
fixed and move the nested family of prefix functionals.

The ordering is post hoc: constructing it requires knowing which placements are covered, hence
the full score vector or its ranks. It is a deterministic traversal for orbit computations and
repeated-placement audits, not an online algorithm, and it does not replace exchangeability in
one-shot conformal prediction, where a single placement is designated as the test; only the
distributional symmetry makes the designated coordinate inherit the orbit-average coverage
probability.

Several coverage levels can be balanced at once, and here the vector-valued lemma of
Section~\ref{sec:steinitz} is what does the work. The sharp one-dimensional $\tfrac12$ bound
does not in general survive at two levels simultaneously: with $N=3$ and levels $k_1=1$,
$k_2=2$, the $t=1$ constraints force the first traversed rank to be $2$, and then either
completion violates $\tfrac12$ at $t=2$, so no ordering of the three placements meets the
scalar bound at both levels, though each level alone admits two.

\begin{proposition}[Simultaneously balanced levels]\label{prop:multilevel}
Fix levels $k_1,\dots,k_d\in\{1,\dots,N\}$ and set
$z_i(s)=\bigl(\ind{\mathrm{rank}(s_i)\le k_\ell}-k_\ell/N\bigr)_{\ell\le d}\in\R^{d}$. Then
$\sum_i z_i(s)=0$ and $\|z_i(s)\|_\infty\le 1$, so the Steinitz lemma supplies one permutation
$\pi$ with
\[
\max_{\ell\le d}\;\Bigl|\frac1t\sum_{j=1}^{t}\ind{\mathrm{rank}(s_{\pi(j)})\le k_\ell}
-\frac{k_\ell}{N}\Bigr|\;\le\;\frac{C_d}{t}
\qquad\text{for every } t=1,\dots,N,
\]
with $C_d$ the Steinitz constant of $(\R^d,\|\cdot\|_\infty)$, independent of $N$.
\end{proposition}

\begin{proof}
Each coordinate of $z_i(s)$ lies in $[-1,1]$. Ranks are a permutation of $\{1,\dots,N\}$, so
$\sum_i\ind{\mathrm{rank}(s_i)\le k_\ell}=k_\ell$ directly, for every
$k_\ell\in\{1,\dots,N\}$, and summing coordinate $\ell$ over $i$ gives $k_\ell-N\cdot
k_\ell/N=0$. Apply the Steinitz lemma in $(\R^d,\|\cdot\|_\infty)$ and divide the prefix bound
by $t$.
\end{proof}

For $k_\ell\le N-1$ the level is the placement-acceptance count of
Proposition~\ref{prop:count}; $k_\ell=N$ is the trivial always-accepted rank level, whose
reading as a conformal threshold needs the $+\infty$ endpoint convention.

A single traversal is therefore calibrated at every selected confidence level at once, at the
price of the dimensional constant in place of the sharp $\tfrac12$.

\section{Balanced sampling: the measure-side counterpart}\label{sec:cubature}

If the pairing of interest is $\langle h,\nu\rangle=\int h\,d\nu$ rather than the permutation
representation, calibration has a measure-side counterpart obtained by reversing the roles of
the two arguments. Calibration holds the empirical score measure $\nu_n$ fixed and chooses a
test: find a threshold $h_q$ with $\langle h_q,\nu_n\rangle=k/n$. The role-reversed problem
holds tests $h_1,\dots,h_d$ and targets $m_1,\dots,m_d$ fixed and chooses a measure: find $\nu$
with $\langle h_j,\nu\rangle=m_j$. (This is an inverse moment problem under the pairing, not
the transpose of an operator: the transpose under the pairing of the moment map
$T(\nu)=(\langle h_j,\nu\rangle)_{j\le d}$ is $T^{*}a=\sum_j a_jh_j$, which lives on the
function side.) Over a finite candidate population with $A_{ji}=h_j(x_i)$ the measure-side
problem is the selection problem $Az=b$, $z\in\{0,1\}^N$, the selected measure being
$\nu_z=\sum_{i\le N}a_iz_i\delta_{x_i}$ with $a_i$ the normalizing or Horvitz--Thompson
weights. Once prescribed inclusion
probabilities and Horvitz--Thompson auxiliary-total constraints are encoded, this becomes the
balanced-sampling problem addressed (exactly when geometry and integrality permit,
approximately otherwise) by the cube method \citep{devilletille2004}. When arbitrary positive
weights are allowed the problem changes and convex geometry answers it exactly. Let $P$ be a
probability measure on $\mathcal X$ and $\phi=(h_1,\dots,h_d):\mathcal X\to\R^d$ measurable
with $\int\|\phi\|\,dP<\infty$. The barycenter $\mu=\int\phi\,dP$ then lies in the convex
hull of $\phi$ on a $P$-full set, so Carath\'eodory applied to the moment map supplies
$m\le d+1$ points $x_1,\dots,x_m$ and weights $w_r>0$ with $\sum_r w_r=1$ and
$\mu=\sum_r w_r\phi(x_r)$: adjoining the constant $1$ to $\phi$ is what preserves total mass
and gives the $d+1$ bound \citep{tchakaloff1957}. In this formulation, balanced sampling and exact
cubature are the measure-side counterpart of threshold calibration.

A designed calibration set sits inside this frame. If designed scores $S_1,\dots,S_n$, taken
distinct, have empirical distribution function within Kolmogorov distance $D$ of a continuous
reference law $\bar F$, and the test score is drawn from $\bar F$ independently of the design,
then $|\bar F(S_{(k)})-k/(n+1)|\le D+1/(n+1)$ and $|\bar F(S_{(k)})-p|<D+(1+p)/n$,
deterministically in the design (the proof is $\hat F_n(S_{(k)})=k/n$ plus
Lemma~\ref{lem:steer}). The midpoint quantile grid $S_i=\bar F^{-1}\bigl((i-\tfrac12)/n\bigr)$, $i=1,\dots,n$,
achieves $D\le 1/(2n)$ for continuous $\bar F$. This is design-based rather
than distribution-free: it suppresses the $n^{-1/2}$ fluctuation of calibration-conditional
coverage at the price of assuming the target law. Two limitations should be stated. The
formulation treats the score values as design variables; realizing such a design through
selectable observations, whose scores are outputs of a predictor and unknown responses, is a
separate problem. And no implication from the reproducing-kernel discrepancy guarantees of
kernel herding or kernel thinning \citep{chen2010,dwivedi2021} to the Kolmogorov discrepancy
required here is established.

\section{The Barber--Pananjady floor}\label{sec:bp}

The two-ingredient structure of the preceding sections reappears in the dependent-data coverage
floor of \citet{barberpananjady2026}. For $S=(S_1,\dots,S_{N})$ they define subvectors
$\Delta^{0}_{k,\tau}(S)$ and $\Delta^{1}_{k,\tau}(S)$ of length $N-\tau$, deleting a block of
$\tau$ entries while keeping $S_{N}$ (respectively rotating $S_k$) in the final slot, and set
\[
\Psi_{k,\tau}(S)=d_{\mathrm{TV}}\bigl(\mathcal L(\Delta^{0}_{k,\tau}(S)),
\mathcal L(\Delta^{1}_{k,\tau}(S))\bigr),
\qquad
\bar\Psi_\tau=\frac1{N}\sum_{k=1}^{N}\Psi_{k,\tau}(S),
\]
with $d_{\mathrm{TV}}(P,Q)=\sup_A|P(A)-Q(A)|$. Their Theorem~5 gives, for a pretrained
memoryless score,
\[
\Prob\{S_{N}\le q_n\}\;\ge\;1-\alpha-\min_{0\le\tau\le n}
\Bigl\{\tfrac{\tau}{N}+\bar\Psi_\tau\Bigr\},
\]
their Proposition~3 bounds the individual switch coefficients by $2\beta(\tau)$ for stationary
$\beta$-mixing data, and, after data processing and averaging, their Corollary~6 inserts this
control into the coverage bound. Their Theorem~8 is a corresponding upper bound with the same
switch-coefficient structure, under almost surely distinct scores; their Theorem~7 shows
tightness up to the factor $(1-\alpha)/4$ when $(1-\alpha)N$ is integral and the data space is
infinite, a finite space carrying an additional cardinality
remainder.\footnote{Result numbering follows the ALT 2026 (PMLR volume 313) version of
\citet{barberpananjady2026}; the arXiv versions number the same results differently.}

The argument separates into an exact insertion--deletion stability statement for empirical
quantile \emph{levels} (deleting $\tau$ entries moves the rank of a fixed threshold by at most
$\tau$, so the level shifts by at most $\tau/N$: order-statistic geometry, controlling levels
rather than values, since empirical quantile values are not stable under small total-variation
perturbations) and a total-variation comparison between switched laws, which is the definition
of the switch coefficient. Both ingredients are linear in their perturbation sizes and no
concentration step appears; by comparison the blocking and empirical-process bound of
\citet{oliveira2024} has square-root dependence on the corresponding block-and-mixing
expression, up to logarithmic factors. This decomposition is thematically parallel to the
discrepancy constructions above, but it is not derived from the contragredient identity.

\section{Three rates}\label{sec:rates}

The shared algebra can blur three quantities. The marginal coverage $k/N$ of
Corollary~\ref{cor:valid} is an exact symmetry identity; its $O(1/n)$ deviation from $p$ is
rank-grid rounding, not estimation error. The calibration-sample-conditional coverage
$\Prob\{S_{n+1}\le q_n\mid S_1,\dots,S_n\}=\bar F(q_n)$, defined when the test score is drawn
from a continuous $\bar F$ independently of the calibration scores, fluctuates around nominal
at the sampling scale: for $S_1,\dots,S_n$ i.i.d.\ from $\bar F$ one has exactly
$\bar F(S_{(k)})\sim\mathrm{Beta}(k,\,n+1-k)$, with standard deviation $O(n^{-1/2})$
\citep{vovk2005}. The design of Section~\ref{sec:cubature} suppresses this at the price of
assuming $\bar F$. Steinitz discrepancies are genuine approximation errors, $O(1/t)$ with
constants from the Steinitz lemma. The ledgers match; the objects differ.

\section{Remarks}\label{sec:remarks}

\begin{remark}[The pinball dual]
Threshold calibration also has an exact convex dual. The conformal threshold $q_n=S_{(k)}$ is
the empirical quantile at level $\gamma=k/n$, not generally at $p$, and it minimises the
pinball loss \citep{koenker1978} at that level, $\sum_i\rho_\gamma(S_i-q)$ with
$\rho_\gamma(r)=\sup_{u\in[\gamma-1,\gamma]}ur$. Fenchel duality gives
\[
\max\Bigl\{\textstyle\sum_i u_iS_i:\ u_i\in[\gamma-1,\gamma],\ \sum_i u_i=0\Bigr\},
\]
and with distinct scores $u_i=\gamma-h_{q_n}(S_i)$ is feasible for every $i$, including at the
threshold, since $\sum_i u_i=n\gamma-k=0$. The dual optimum is a balanced signed measure on the
scores: calibration's exact optimization dual, though not itself a point-generation algorithm.
\end{remark}

\begin{remark}[Herding]\label{rmk:herding}
Herding \citep{welling2009} generates points by the recursion $w_t=w_{t-1}+\mu-\phi(x_t)$,
$x_t=\arg\max_x\langle w_{t-1},\phi(x)\rangle$, and attains $O(1/n)$ feature-mean discrepancy
conditional on bounded weights; the identification with Frank--Wolfe \citep{bach2012} shows the
fast rate needs interiority-type conditions, with \citet{chen2010} treating the kernel case.
\citet{harvey2014} introduce a Permutation Problem, each point used once, that is exactly the
balanced-permutation problem of Section~\ref{sec:steinitz}, and show that solving it solves
herding over the finite population; generic herding runs with replacement, typically on
nonfinite domains \citep{welling2009,chen2010}, and on the pairing of
Section~\ref{sec:cubature} it is an iterative with-replacement method for approximate moment
matching, or equal-weight cubature; this is related to, but not identical with, balanced
sampling in the Deville--Till\'e sense, which fixes inclusion probabilities and keeps the
$0$--$1$ without-replacement structure. The quadrature triangle is older on that side: optimally-weighted
herding is Bayesian quadrature \citep{huszar2012}, and \citet{snellgriffiths2025} recently
recast conformal risk and threshold calculations through Bayesian quadrature of the loss
quantile function.
\end{remark}

\begin{remark}[Limits]
The passage from orbit identity to one-shot coverage is exchangeability's, or its mixing
surrogate's in Section~\ref{sec:bp}; no bookkeeping supplies it. Nor does control of thresholds
and labels create information: it re-levels coverage while conditional sharpness remains a
property of the scores \citep{cotton_marginally}. Whether placement balancing can be aimed
conditionally, selecting or reweighting calibration slots to balance the rank count given the
input, is open; localized and weighted conformal methods are the nearest existing steps.
\end{remark}

\bibliographystyle{plainnat}
\bibliography{../notes}

\end{document}
