\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{microtype}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{remark}{Remark}

\title{The Width of the Conformal Fan:\\
Dependence and the Variance of Realized Coverage}
\author{Peter Cotton}
\date{Draft --- companion note}

\begin{document}
\maketitle

\begin{abstract}
Fix a split-conformal calibration set and ask how often a fresh, \emph{independent} draw from the
score marginal falls under the threshold $U_{(k)}$. That realized coverage $c=U_{(k)}$ is a random
variable; for independent calibration scores it follows the $\mathrm{Beta}(k,n-k+1)$ law, variance
about $\alpha(1-\alpha)/n$, the classical fan. We study how dependence among the calibration scores
reshapes it. The organizing identity is that $c$ is a functional of the count process
alone, so its variance is governed by the count's covariance kernel, and negatively associated scores
have a less variable count (Lemma~\ref{lem:count}). Three results follow. A leading-order fan
coefficient: in a Bahadur regime $\operatorname{Var}(c)=n^{-2}\operatorname{Var}(N(1-\alpha))+o(1/n)$,
so the average pairwise covariance of the sub-threshold indicators sets the leading constant. An exact
aggregate bound: summed over all levels $k$, negative association never widens the fan, by a
convex-order contraction of the count against a concave functional and majorization. And, for
equicorrelated normals, the single level is exact and monotone in the correlation,
$\operatorname{Var}(Z_{(k)})=v_k+\rho(1-v_k)$, dispersively so via strong log-concavity and Tweedie's
formula. We keep $c$ (coverage against an independent test draw) distinct from conformal conditional
coverage with an exchangeable test point: under dependence the \emph{mean} of $c$ is not pinned, and
the positive-dependence decomposition we give is an exact identity that does not by itself order the
fan against the independent one. The general single-level negative-association contraction is left as
a conjecture, supported numerically. The governing sign is the one the companion note attaches to the
finite de Finetti measure.
\end{abstract}

\section{Setup}

Split conformal returns a set with marginal coverage $1-\alpha$. Fix the calibration set; the
coverage you then realize on fresh test points is a random variable, and it is not $1-\alpha$ on the
nose. Write the scores on the probability-integral scale, $U_i=F(S_i)$. We assume throughout that the
score marginal $F$ is continuous, so each $U_i\sim\mathrm{Unif}(0,1)$; if $F$ has atoms, read $U_i$ as
the randomised PIT $F(S_i^-)+V_i\,(F(S_i)-F(S_i^-))$ with $V_i\sim\mathrm{Unif}(0,1)$ independent,
which restores exact uniformity. With $k=\lceil(n+1)(1-\alpha)\rceil$ the conformal threshold is the
order statistic $U_{(k)}$, and the calibration-conditional coverage against an independent test draw
from the marginal is
\begin{equation}
c=U_{(k)} .
\end{equation}
For \emph{independent} calibration scores this is the classical result \citep{vovk2012,marques2024}:
$c\sim\mathrm{Beta}(k,n-k+1)$, with
\begin{equation}\label{eq:betavar}
\mathbb E[c]=\frac{k}{n+1},\qquad \operatorname{Var}(c)=\frac{k(n-k+1)}{(n+1)^2(n+2)}\;\approx\;
\frac{\alpha(1-\alpha)}{n}.
\end{equation}
This is the fan: the per-dataset coverage scatters around $1-\alpha$ with variance of order $1/n$
(standard deviation of order $n^{-1/2}$). The question of this note is what the dependence among the
calibration scores does to it.

Two cautions fix the object of study. First, $c=U_{(k)}$ is coverage against a test draw
\emph{independent} of the calibration set; it is not the conformal conditional coverage
$\mathbb P(U_{n+1}\le U_{(k)}\mid U_1,\dots,U_n)$ for a test point exchangeable with, and dependent
on, the calibration scores, which need not equal $U_{(k)}$. The marginal conformal guarantee comes
from the rank symmetry of $U_{n+1}$ among the $n+1$ scores and is not what controls $c$. Second, the
mean $\mathbb E[c]=k/(n+1)$ is an \emph{independence} statement, not a consequence of uniform
marginals or exchangeability: if $U_1=\dots=U_n=V\sim\mathrm{Unif}(0,1)$ then $U_{(k)}=V$ and
$\mathbb E[c]=\tfrac12$ for every $k$. So under dependence both the mean and the variance of $c$
respond; this note studies the variance, and flags the mean where it matters. The companion notes
treat the orthogonal, within-dataset axis \citep{cotton_marginally,cotton_signed}; here every
statement is about the across-dataset fluctuation of $c$. The results separate by type:
Theorems~\ref{thm:pos} and \ref{thm:agg} and Propositions~\ref{prop:path}--\ref{prop:gauss} are
finite-sample exact, Theorem~\ref{thm:coeff} is asymptotic and conditional on a Bahadur
representation, and the general per-level negative-association contraction (Conjecture~\ref{conj:disp})
is open.

The whole problem passes through the count process $N(u)=\sum_{i=1}^n\mathbf 1\{U_i\le u\}$, because
$U_{(k)}=\int_0^1\mathbf 1\{N(u)<k\}\,\mathrm du$, so
\begin{equation}\label{eq:reduction}
\operatorname{Var}(c)=\int_0^1\!\!\int_0^1\operatorname{Cov}\big(\mathbf 1\{N(u)<k\},\,
\mathbf 1\{N(v)<k\}\big)\,\mathrm du\,\mathrm dv,
\end{equation}
a functional of the law of the count process alone --- equivalently, by Sklar's theorem, of the score
\emph{copula}, since the uniform marginal has factored out (we return to this in
Remark~\ref{rmk:copula}). Two facts about $N$ drive everything.

\begin{lemma}[Count variance]\label{lem:count}
If the scores are negatively associated \citep{joagdev1983}, then for every $u$,
$\operatorname{Var}(N(u))\le n\,u(1-u)$, with equality iff the sub-threshold indicators are pairwise
uncorrelated.
\end{lemma}
\begin{proof}
The indicators $\mathbf 1\{U_i\le u\}$ are functions of single distinct coordinates, all nondecreasing
in the same direction, so negative association is preserved and they are pairwise non-positively
correlated. So
$\operatorname{Var}(N(u))=\sum_i\operatorname{Var}+\sum_{i\ne j}\operatorname{Cov}\le
\sum_i u(1-u)=n\,u(1-u)$.
\end{proof}

\section{The fan coefficient}

Let $p=1-\alpha$ and $b_i=\mathbf 1\{U_i\le p\}$, so $N(p)=\sum_i b_i$. (We reserve $\xi$ for the
i.i.d.\ normals of Proposition~\ref{prop:gauss}.) The representation is on the uniform scale, where the
quantile density is $1/f(p)=1$, so the usual $1/f$ factor is absent.

\begin{theorem}[Fan coefficient]\label{thm:coeff}
Suppose $F$ is continuous (or a randomised PIT is used), $k/n\to p\in(0,1)$, and the scores satisfy a
Bahadur representation $U_{(k)}-p=p-N(p)/n+r_n$ with the remainder negligible in $L^2$ relative to
$n^{-1}\operatorname{Var}(N(p))^{1/2}$. Then
\begin{equation}
\operatorname{Var}(c)=\frac1{n^2}\operatorname{Var}(N(p))+o\!\Big(\frac{\operatorname{Var}(N(p))}{n^2}\Big).
\end{equation}
If moreover $\operatorname{Var}(N(p))\asymp n$, writing it as
$np(1-p)+n(n-1)\bar c$ with $\bar c=\tfrac1{n(n-1)}\sum_{i\ne j}\operatorname{Cov}(b_i,b_j)$,
\begin{equation}
\operatorname{Var}(c)=\frac{p(1-p)}{n}\,R\,(1+o(1)),\qquad R=1+\frac{n-1}{p(1-p)}\,\bar c .
\end{equation}
\end{theorem}

So the leading fan coefficient is the count variance at the operating quantile. Under
independence $\bar c=0$, $R=1$, recovering \eqref{eq:betavar}; negative association makes $\bar c<0$
and $R<1$, narrowing the fan. Two caveats the statement respects. The multiplicative form needs
$\operatorname{Var}(N(p))\asymp n$: under strong negative dependence the count variance can be
$o(n)$ (zero at the floor), and then only the additive form is safe. And the Bahadur regime itself
fails under a fixed positive latent (de Finetti) dependence, where $N(p)/n$ has a non-degenerate
random limit $Q([0,p])$ and $U_{(k)}$ converges to a random conditional quantile rather than
linearizing at $p$; the next section handles that regime exactly and separately.

\section{Positive dependence: an exact decomposition}

When the scores are extendable, drawn from an infinitely exchangeable sequence, de Finetti
\citep{definetti1937,kerns2006} makes them conditionally independent given a latent directing measure
$Q$. The law of total variance then gives an exact decomposition.

\begin{theorem}[Positive-dependence decomposition]\label{thm:pos}
For an extendable exchangeable calibration sequence with directing measure $Q$,
\begin{equation}
\operatorname{Var}(c)=\underbrace{\mathbb E_Q\!\big[\operatorname{Var}(U_{(k)}\mid Q)\big]}
_{\text{average within-}Q\text{ order-stat variance}}
+\underbrace{\operatorname{Var}_Q\!\big(m_k(Q)\big)}_{\text{between-}Q\text{ term}},
\qquad m_k(Q)=\mathbb E[U_{(k)}\mid Q].
\end{equation}
The between-$Q$ term is non-negative and vanishes under independence.
\end{theorem}

The decomposition is exact and needs no asymptotics, but note what conditioning gives: $U_i\mid Q$ are
i.i.d.\ $Q$, so $U_{(k)}\mid Q$ is the $k$-th order statistic of $Q$ on the marginal scale, \emph{not}
a uniform $\mathrm{Beta}$ variable, and its mean $m_k(Q)$ moves with $Q$. The within-$Q$ term is thus
not the independent fan \eqref{eq:betavar}, and the identity by itself does \emph{not} prove
$\operatorname{Var}(c)$ exceeds the independent value: that comparison would need a separate theorem
relating the average within-$Q$ order-stat variance to the uniform fan. What the identity does isolate
is the between-$Q$ term, one signature of positive dependence (it also shifts the within-$Q$ law from
uniform to $Q$): a per-dataset latent shifts the whole
calibration sample, and the realized coverage rides with it. Empirically the net effect is a wider
fan, e.g.\ a one-factor Gaussian copula at correlation $0.2$ gives $\operatorname{Var}(c)\approx 3.3$
times the independent value (and the comonotone limit $U_1=\dots=U_n$ gives $\tfrac1{12}$, far above
it), but that widening is observed, not proved by Theorem~\ref{thm:pos} alone.

\section{Negative dependence narrows the fan}

The negative regime is the one a genuine prior cannot reach. By \citet{kerns2006} a finite
exchangeable sequence with $\rho<0$ has a \emph{signed} directing measure, so the clean conditioning
of Theorem~\ref{thm:pos} is unavailable: there is no $Q$ to condition on, and the between-dataset
term would be a negative quasi-variance. That is precisely the regime where the fan contracts.

\begin{theorem}[Endpoints]\label{thm:ends}
Independent scores give $c\sim\mathrm{Beta}(k,n-k+1)$, the benchmark fan. At the exchangeable
pairwise-correlation floor, take the discrete rank idealization in which the calibration scores are a
uniformly random permutation of the grid $\{1/(n+1),\dots,n/(n+1)\}$ (sampling without replacement,
pairwise correlation $-1/(n-1)$, the smallest attainable for an exchangeable vector): the order
statistic is deterministic, $U_{(k)}=k/(n+1)$ a.s., so $\operatorname{Var}(c)=0$.
\end{theorem}

Independence is the benchmark, not the maximum: positive dependence can widen the fan well past it
(the comonotone limit reaches $\tfrac1{12}$). The claim of this note is the other direction, that
\emph{negative} association does not widen it, with independence the (conjectural) upper endpoint of
the negative-dependence range. Two model cautions on the floor. The grid permutation has
\emph{discrete}-uniform marginals, not the continuous $\mathrm{Unif}(0,1)$ of $U_i=F(S_i)$, so it sits
in the model only if one allows discrete or rank scores; the continuous-marginal analogue is a Latin
hypercube (one point per stratum, randomly permuted), whose fan is small, of order $1/n^2$, but not
exactly zero. And the endpoints agree with Theorem~\ref{thm:coeff} at $R=1$ and $R=0$. The two are
joined by an exact path.

\begin{proposition}[An exact variance interpolation]\label{prop:path}
Let the calibration law be, with probability $1-t$, $n$ independent uniforms, and with probability
$t$, a contest (a uniformly random permutation of the grid $\{1/(n+1),\dots,n/(n+1)\}$). Both
components yield realized coverage with the same mean $k/(n+1)$, the contest with zero variance, so by
the law of total variance the between-component term vanishes and
\begin{equation}
\operatorname{Var}(c)=(1-t)\,\frac{k(n-k+1)}{(n+1)^2(n+2)} .
\end{equation}
\end{proposition}

\begin{proof}
Condition on the component. $\mathbb E[c\mid\text{iid}]=\mathbb E[c\mid\text{contest}]=k/(n+1)$, so
$\operatorname{Var}(\mathbb E[c\mid\cdot])=0$, and $\mathbb E[\operatorname{Var}(c\mid\cdot)]=
(1-t)\,\tfrac{k(n-k+1)}{(n+1)^2(n+2)}$.
\end{proof}

So the fan collapses linearly to zero as $t\to1$. Two honest caveats. The mixture's pairwise
correlation is $\operatorname{Corr}(U_i,U_j)=-t/(n+1-2t)$, not $-t/(n-1)$: the grid component
contributes $\operatorname{Cov}=-1/(12(n+1))$ but mixing changes the marginal variance to
$(n+1-2t)/(12(n+1))$; the correlation reaches the floor $-1/(n-1)$ only at $t=1$. And the mixture's
one-dimensional marginal is neither continuous uniform nor discrete-grid uniform, so this is an exact
\emph{variance} interpolation, not a fixed-marginal copula path.

So we can prove the contraction exactly at both endpoints and along this whole interpolating family,
and to leading order everywhere (Theorem~\ref{thm:coeff}). For an \emph{arbitrary} negatively
associated law the contraction also holds once it is summed over all levels.

\begin{theorem}[Aggregate fan bound]\label{thm:agg}
For negatively associated scores with uniform marginals,
\begin{equation}
\sum_{k=1}^n\operatorname{Var}(U_{(k)})\;\le\;\sum_{k=1}^n\frac{k(n-k+1)}{(n+1)^2(n+2)},
\end{equation}
the independent value, with equality holding under independence. Summed over all target levels,
the fan is never widened by negative association.
\end{theorem}

\begin{proof}
The order statistics are a permutation of the sample, so $\sum_k U_{(k)}^2=\sum_i U_i^2$ and, under
uniform marginals, $\sum_k\mathbb E[U_{(k)}^2]=n\,\mathbb E[U_1^2]=n/3$ for every such law; hence
$\sum_k\operatorname{Var}(U_{(k)})=\tfrac n3-\sum_k(\mathbb E\,U_{(k)})^2$. For the top $j$ statistics,
$\sum_{k>n-j}U_{(k)}=\int_0^1\min\!\big(j,n-N(u)\big)\,\mathrm du$. The function $\phi(m)=\max(-j,m-n)$
is convex and \emph{nondecreasing}, which is exactly the class for which \citet{shao2000} compares a
negatively associated sum to its independent counterpart, $\mathbb E\phi(N(u))\le\mathbb E\phi(\mathrm{Bin}(n,u))$.
(Because the means agree, $\mathbb E N(u)=nu=\mathbb E\,\mathrm{Bin}$, one may add any affine term, so
the comparison in fact extends to all convex $\phi$, the full convex order; the monotone-convex form
already suffices here.) Since $\min(j,n-m)=-\phi(m)$, this gives
$\mathbb E\min(j,n-N(u))\ge\mathbb E\min(j,n-\mathrm{Bin}(n,u))$; integrating,
\begin{equation}\label{eq:topj}
\sum_{k>n-j}\mathbb E\,U_{(k)}\;\ge\;\sum_{k>n-j}\tfrac{k}{n+1}\qquad\text{for every }j=1,\dots,n,
\end{equation}
the right side being the independent profile's top-$j$ sum. Both profiles are increasing and have the
same total, $\sum_k\mathbb E\,U_{(k)}=\sum_i\mathbb E\,U_i=n/2=\sum_k\tfrac k{n+1}$ (the case $j=n$ of
\eqref{eq:topj}, with equality). For two increasing vectors with equal sum, ``every top-$j$ partial
sum of one dominates that of the other'' is exactly the definition of majorization; so
$(\mathbb E\,U_{(1)},\dots,\mathbb E\,U_{(n)})$ majorizes $(\tfrac1{n+1},\dots,\tfrac n{n+1})$. The map
$x\mapsto\sum_k x_k^2$ is Schur-convex, hence larger on the majorizing vector,
$\sum_k(\mathbb E\,U_{(k)})^2\ge\sum_k(\tfrac k{n+1})^2$, and the displayed identity for
$\sum_k\operatorname{Var}(U_{(k)})$ finishes it.
\end{proof}

The proof uses only identical uniform marginals and the convex-order domination
$N(u)\le_{\mathrm{cx}}\mathrm{Bin}(n,u)$, not exchangeability as such. Equality holds under
independence; that equality \emph{forces} independence is not established here and would need a
separate tightness lemma. The per-level refinement, for the single conformal $k$, remains open, and
the obstruction is exactly
why: the step $\mathbf 1\{N(u)<k\}$ behind a single $\operatorname{Var}(U_{(k)})$ is neither concave
nor convex, so the convex order does not sign it, whereas the smooth aggregate functional
$\min(j,n-N)$ is concave and it does.

\begin{conjecture}[Per-level contraction]\label{conj:disp}
If the calibration scores are negatively associated then $\operatorname{Var}(c)$ is at most the
independent value \eqref{eq:betavar} for each $k$ individually.
\end{conjecture}

In the canonical continuous family the per-level statement is exact, and more, it is monotone in the
correlation.

\begin{proposition}[Exact per-level law for equicorrelated normals]\label{prop:gauss}
Let $(Z_1,\dots,Z_n)$ be equicorrelated standard normal with correlation
$\rho\in[-\tfrac1{n-1},1]$, and let $v_k=\operatorname{Var}(\xi_{(k)})$ be the variance of the $k$-th
order statistic of $n$ i.i.d.\ standard normals. Then, for every $k$,
\begin{equation}
\operatorname{Var}(Z_{(k)})=v_k+\rho\,(1-v_k).
\end{equation}
As $v_k<1$ for $n\ge 2$, this is strictly increasing in $\rho$: negative correlation strictly shrinks
the variance of every order statistic, positive correlation inflates it.
\end{proposition}

\begin{proof}
Write $Z_i=\bar Z+(Z_i-\bar Z)$. For equicorrelated normals $\bar Z\sim N(0,s)$,
$s=(1+(n-1)\rho)/n$, is independent of the centred vector, and
$\operatorname{Cov}(Z_i-\bar Z,Z_j-\bar Z)=(1-\rho)(\delta_{ij}-1/n)$, exactly $1-\rho$ times the
covariance of $(\xi_i-\bar\xi)$ for i.i.d.\ standard normals. Hence
$Z_{(k)}\stackrel{d}{=}\sqrt{s}\,\eta+\sqrt{1-\rho}\,(\xi-\bar\xi)_{(k)}$ with $\eta\sim N(0,1)$
independent of the second term, so
$\operatorname{Var}(Z_{(k)})=s+(1-\rho)\operatorname{Var}\big((\xi-\bar\xi)_{(k)}\big)$. Both terms
are affine in $\rho$; matching $v_k$ at $\rho=0$ forces
$\operatorname{Var}((\xi-\bar\xi)_{(k)})=v_k-1/n$, and substituting gives $v_k+\rho(1-v_k)$.
\end{proof}

\begin{remark}
This is the per-level contraction, exact and monotone, on the Gaussian scale, unifying the two sides:
the single line $v_k+\rho(1-v_k)$ runs from inflation at $\rho>0$ through the independent value at
$\rho=0$ to the floor at $\rho=-1/(n-1)$. Under the Gaussian copula $U_i=\Phi(Z_i)$, which restores
uniform marginals, $\operatorname{Var}(U_{(k)})$ is numerically monotone in $\rho$ and below the
independent value throughout $\rho<0$ (\texttt{check\_fan.py}). The transfer to the uniform scale is
a dispersive-order statement, and it is half rigorous. Decompose along the common mean,
$Z_{(k)}=\sqrt{1-\rho}\,D+\bar Z$ with $D=(\xi-\bar\xi)_{(k)}$ and $\bar Z\sim N(0,s)$ independent.
$D$ is log-concave, being the order statistic of a vector with log-concave (Gaussian) joint density
\citep{prekopa1973}; and convolution with a log-concave density increases the dispersive order
\citep{lewis1981}, so adding the Gaussian shift $\bar Z$ disperses, monotonically in $s$. The residual
is that this outweighs the simultaneous $\sqrt{1-\rho}$ shrinkage of $D$ as $\rho$ grows, a one-parameter
density-quantile monotonicity under the variance budget $\tfrac{n-1}n(1-\rho)+s=1$, which no
off-the-shelf convolution theorem delivers but which we now settle.

Tweedie's formula sharpens that residual to a single posterior-variance bound. Reading
$Y:=Z_{(k)}=aD+b\eta$ ($a=\sqrt{1-\rho}$, $b=\sqrt s$) as Gaussian signal extraction of the
log-concave signal $D$, the slope of the posterior mean equals the scaled posterior variance,
$\frac{d}{dq}\mathbb E[D\mid Y=q]=a\,\operatorname{Var}(D\mid Y=q)/b^2$
\citep{efron2011,hansen2026}. The dispersive monotonicity is equivalent to this slope being at most
$(n-1)a/n$, so, using $\tfrac{n-1}n a^2+b^2=1$, the entire transfer reduces to
\begin{equation}
\operatorname{Var}(D\mid Y=q)\;\le\;\frac{n-1}{n}\,b^2\qquad\text{for all }q,
\end{equation}
a posterior-variance bound for log-concave Gaussian denoising, which the next lemma settles. The
score-driven reading of the same identity \citep{hansen2026} is the econometric form of modelling the
conditional spread, which is exactly what closes the information gap a conformal wrapper leaves
\citep{cotton_marginally}.
\end{remark}

\begin{lemma}[Strong log-concavity of $D$]\label{lem:slc}
$D=(\xi-\bar\xi)_{(k)}$ is $\tfrac{n}{n-1}$-strongly log-concave, $(\log g_D)''\le -n/(n-1)$.
\end{lemma}

\begin{proof}
The centred vector $w=\xi-\bar\xi$ has covariance $I-\tfrac1n J$, an orthogonal projection, so on the
hyperplane $\{\sum_i w_i=0\}$ it is a \emph{standard} Gaussian: its $-\log$ density is $\tfrac12\|w\|^2$,
with Hessian the identity. The law of the ordered vector is this density restricted to the convex cone
$\{w_1\le\dots\le w_n\}$. The cone enters only through the indicator of a convex set, a log-concave
factor; writing the restricted density as $\exp(-\tfrac12 w^\top(I-\tfrac1nJ)^{+}w-\iota_C(w))$ with
$\iota_C$ the convex indicator of the cone, the Hessian of the potential is
$(I-\tfrac1nJ)^{+}+\nabla^2\iota_C\succeq(I-\tfrac1nJ)^{+}$. So the strong-log-concavity constant is
set by the Gaussian quadratic alone; the cone can only raise it, never lower it, and in particular
does not interfere through its boundary. Strong log-concavity passes to marginals
\citep{prekopa1973,brascamplieb1976,saumard2014}, the marginal of a coordinate being strongly
log-concave to the reciprocal of that coordinate's reference variance; for $w_{(k)}$ this is
$\tfrac{n-1}n$ (the diagonal of $I-\tfrac1nJ$), giving constant at least $\tfrac n{n-1}$ (the
parameter bookkeeping is in the Appendix).
\end{proof}

\begin{proposition}[Gaussian-scale dispersive monotonicity]\label{prop:dispmon}
For equicorrelated standard normals with $\rho\in(-\tfrac1{n-1},1)$, $Z_{(k)}$ is increasing in the
dispersive order as $\rho$ increases.
\end{proposition}

\begin{proof}
For a smooth one-parameter family the quantile evolves by $\partial_\rho q_\rho(u)=
\mathbb E[\dot Y\mid Y=q_\rho(u)]$ with $\dot Y=\partial_\rho Z_{(k)}$, and dispersive monotonicity is
exactly $\partial_\rho q_\rho(u)$ increasing in $u$, i.e.\ increasing in $q$. Writing
$Y=aD+b\eta$ and using $a\,\mathbb E[D\mid Y]+b\,\mathbb E[\eta\mid Y]=Y$ together with the budget
$\tfrac{n-1}na^2+b^2=1$, this reduces to $m'(q)\le (n-1)a/n$, where $m(q)=\mathbb E[D\mid Y=q]$ (the
computation of $M'(q)$ is in the Appendix).
Tweedie's identity for Gaussian signal extraction gives $m'(q)=a\operatorname{Var}(D\mid Y=q)/b^2$.
The posterior of $D$ given $Y$ has $-\log$ density with Hessian
$(-\log g_D)''+a^2/b^2\ge \tfrac n{n-1}+\tfrac{a^2}{b^2}=\tfrac{n}{(n-1)b^2}$, the last equality by the
budget and Lemma~\ref{lem:slc}; the Brascamp--Lieb variance inequality \citep{brascamplieb1976} then
gives $\operatorname{Var}(D\mid Y=q)\le (n-1)b^2/n$, hence $m'(q)\le (n-1)a/n$. Equivalently the
quantile density $q_\rho'(u)=1/f_\rho(q_\rho(u))$ increases in $\rho$ at each level (the density at each
quantile decreases), which is the dispersive monotonicity. At the open-interval ends the statement is
read in the limit: $b\to0$ at $\rho=-1/(n-1)$ (the degenerate floor, $Z_{(k)}\to\sqrt{n/(n-1)}\,D$) and
$a\to0$ at $\rho=1$ ($Z_{(k)}\to\eta$).
\end{proof}

This settles the Gaussian-scale per-level statement in the strong (dispersive, not merely variance)
sense. The one step left for the uniform-marginal conjecture under the Gaussian copula is the passage
through $\Phi$: being S-shaped, $\Phi$ is neither convex nor concave, so it need not transmit the
dispersive order, and $\operatorname{Var}(U_{(k)})=\operatorname{Var}(\Phi(Z_{(k)}))$ monotone in
$\rho$ is confirmed numerically (\texttt{check\_fan.py}) but not by a transfer theorem.

The Gaussian case fixes the copula and varies $k$; the opposite slice, the top level $k=n$ for
\emph{any} negatively associated sample, is also exactly tractable and is where the conjecture becomes
one-dimensional. There $U_{(n)}\le u$ iff every score lies below $u$, so the law of $U_{(n)}$ is the
copula diagonal $\delta(u)=C(u,\dots,u)=\mathbb P(N(u)=n)$, and
$\operatorname{Var}(U_{(n)})=\int_0^1\!\!\int_0^1[\delta(u\wedge v)-\delta(u)\delta(v)]\,du\,dv$.
Negative association enters only as the negative-orthant bound $\delta(u)\le u^n$
\citep{joagdev1983} --- equivalently the convex-order image of the discretely convex indicator
$\mathbf 1\{N=n\}$. Writing the deficit $g(u)=u^n-\delta(u)\ge 0$ and $\mu^\ast=n/(n+1)=\mathbb
E[U_{(n)}^{\mathrm{iid}}]$, the variance has the exact closed form
\begin{equation}\label{eq:maxdiag}
\operatorname{Var}(U_{(n)})=\operatorname{Var}^{\mathrm{iid}}(U_{(n)})
+2\int_0^1(u-\mu^\ast)\,g(u)\,du-\Big(\int_0^1 g\Big)^{\!2},
\end{equation}
so the per-level conjecture at $k=n$ is \emph{equivalent} to a centroid inequality on the deficit,
\begin{equation}\label{eq:centroid}
\frac{\int_0^1 u\,g(u)\,du}{\int_0^1 g(u)\,du}\;\le\;\mu^\ast+\tfrac12\!\int_0^1 g\,:
\end{equation}
the mass by which the diagonal falls short of independence must sit, on average, at most a half-deficit
above the independent-maximum mean. This is \emph{not} a consequence of the diagonal alone. A linear
program over the functions meeting every condition a copula diagonal must satisfy (monotone, $0\to1$,
$\delta\le u$, $n$-Lipschitz, Fr\'echet lower bound; \citealp{nelsen2006}) together with the negative-orthant
bound $\delta\le u^n$ returns a diagonal that violates \eqref{eq:centroid} (\texttt{check\_fan.py}).
The single level is therefore irreducibly a property of the off-diagonal negative-orthant probabilities
across thresholds, not of the diagonal that determines $U_{(n)}$'s law: exactly the content the
aggregate convex-order bound uses in bulk and that a one-level statement cannot read off. It is the same
obstruction that blocks localizing Theorem~\ref{thm:agg} to a single $k$, now in closed form, and every
negatively associated structure tested satisfies \eqref{eq:centroid} with margin.

The same convex-order contraction $N(u)\le_{\mathrm{cx}}\mathrm{Bin}(n,u)$ that drives the aggregate
bound is available per level, but turning it into the single-$k$ statement needs the tail
$\mathbb P(N(u)\ge k)$ to cross the binomial tail once in $u$, which the total positivity of the
binomial tail in $(k,u)$ \citep{karlin1968} makes plausible, followed by a
single-crossing-implies-less-spread step. We claim only the variance statement, and deliberately not
the stronger dispersive ordering: an order statistic of a dependent sample need not be dispersively
smaller than the independent one even at fixed marginals \citep{panja2022}, so the result is specific
to the variance and is not a corollary of a dispersive theorem. The natural sufficient condition on
the dependence is negative supermodular dependence \citep{christofides2004}, which all the structures
below satisfy. The table is illustrative, each row a single Monte Carlo estimate ($6\times10^4$ draws,
so the variance entries carry a relative Monte Carlo error of a few percent), fully reproduced by
\texttt{check\_fan.py} ($n=20$, $\alpha=0.1$, independent fan variance $3.92\times10^{-3}$); it is
evidence for the conjecture, not a proof.

\begin{center}
\begin{tabular}{lcccc}
\hline
calibration dependence & $\mathbb E[c]$ & $\operatorname{Var}(c)$ & ratio to fan & sign \\
\hline
independent ($\rho=0$)                 & $0.904$ & $3.92\times10^{-3}$ & $1.00$ & --- \\
exchangeable, $\rho=-1/(n-1)$          & $0.923$ & $1.16\times10^{-3}$ & $0.29$ & neg.\ assoc. \\
MA(1), lag-1 corr $-\tfrac12$          & $0.912$ & $2.71\times10^{-3}$ & $0.69$ & neg.\ assoc. \\
random negatively-paired blocks        & $0.909$ & $3.14\times10^{-3}$ & $0.80$ & neg.\ assoc. \\
contest / without replacement          & $0.905$ & $0$                  & $0.00$ & floor \\
one-factor, $\rho=+0.2$                & $0.863$ & $1.33\times10^{-2}$ & $3.41$ & positive \\
\hline
\end{tabular}
\end{center}

Every negatively associated structure, exchangeable or not, comes in under the independent fan; the
positive control widens it. The mean of $c$ also responds to dependence, drifting up under negative
and down under positive (to $0.863$ in the one-factor row), confirming it is not pinned; the variance
is what we track here. A wider search agrees:
over exchangeable, moving-average, block, and randomly-generated non-exchangeable negatively-correlated
Gaussian copulas at $n\le 8$ and every $k$, the ratio to the independent fan never exceeds one
(it ranges from about $0.30$ at the floor to $0.80$ for weak dependence), with no counterexample. In all
of these the order-statistic CDFs $\mathbb P(N(u)\ge k)$ cross the independent ones exactly once, the
single crossing the remark below uses.

\begin{remark}[A closed form, and what an exact proof must dominate]
The variance is an explicit functional of $g(u)=\mathbb P(N(u)\ge k)$,
\begin{equation}
\operatorname{Var}(c)=2\int_0^1(1-u)\,g(u)\,\mathrm du-\Big(\int_0^1 g(u)\,\mathrm du\Big)^{\!2},
\end{equation}
so writing $\delta=g-g^{\mathrm{iid}}$, $A=\int_0^1 g^{\mathrm{iid}}=1-\mathbb E[c^{\mathrm{iid}}]$,
and $\Delta=\int_0^1\delta$,
\begin{equation}
\operatorname{Var}(c)-\operatorname{Var}(c^{\mathrm{iid}})
=2\int_0^1\big[(1-u)-A\big]\,\delta(u)\,\mathrm du-\Delta^2 .
\end{equation}
The $-\Delta^2$ term is non-positive. The weight $(1-u)-A$ changes sign at
$u=\mathbb E[c^{\mathrm{iid}}]=k/(n+1)$. Empirically (across every structure in the table, and in the
wider search below) $\delta$ changes sign exactly once, near $u\approx k/n$; we emphasize this is an
\emph{observed} single-crossing, not a consequence of Lemma~\ref{lem:count}, since count-variance
contraction alone does not force one-crossing of the tail curve $g$. \emph{Granting} that single
crossing, the two sign changes agree up to the conformal offset between $k/n$ and $k/(n+1)$, a band of
width $O(1/n)$ on which the integrand is positive and elsewhere negative, so the identity gives the
contraction up to that $O(1/n)$ boundary term, a second route to the leading order of
Theorem~\ref{thm:coeff}, and isolates what a finite-sample proof must control: the boundary remainder
against the slack $\Delta^2$.
\end{remark}

\begin{remark}[Why the closed form does not close it, and the lemma that would]
The closed form is, by itself, circular. Substituting the moment identities
$\int u\,\delta=\tfrac12(\mathbb E[c_{\mathrm{iid}}^2]-\mathbb E[c^2])$ and
$\int\delta=\mathbb E[c_{\mathrm{iid}}]-\mathbb E[c]$ into any elementary rearrangement of the
difference, single-crossing splits, Abel summation, the exact first-order perturbation, collapses it
back to $\operatorname{Var}(c)-\operatorname{Var}(c_{\mathrm{iid}})$ with the auxiliary crossing
point cancelling: these structural facts carry no magnitude. Nor is the per-level convex order of
Lemma-\ref{lem:count} enough on its own. Because $\operatorname{Var}(c)$ depends only on the single
curve $g(u)=\mathbb P(N(u)\ge k)$, one might hope $N(u)\le_{\mathrm{cx}}\mathrm{Bin}(n,u)$ at each
$u$ suffices; it does not, since a curve dominated pointwise can have larger variance (place an atom
at the boundary), and such curves are excluded only because they are not realizable by a genuine
sample. The content is therefore an extremal statement about the realizable curves, not their
marginals:
\begin{quote}
\emph{Open lemma.} Among all $g(u)=\mathbb P(N(u)\ge k)$ realizable by a negatively associated uniform
sample, the independent (binomial) curve maximizes $\operatorname{Var}(c)$.
\end{quote}
This is what a total-positivity or coupling argument would have to deliver. Proposition~\ref{prop:path}
and the contest endpoint are the cases settled by hand.
\end{remark}

\begin{remark}[The signed measure as a negative between-term]
Theorem~\ref{thm:pos} reads the fan as within-dataset variance plus a non-negative between-dataset
variance. In the signed-de-Finetti picture of the companion note, the negative corner is exactly
where that between-dataset object turns negative, a Feynman--Wigner quasi-variance that subtracts from
the within-dataset fan rather than adding to it, reaching the full cancellation $R=0$ at the contest
floor. This is heuristic, since a signed measure has no honest conditional variance, but it is the
right cartoon: the same sign that decides conditional adaptivity in \citet{cotton_signed} decides
whether the fan is inflated or cancelled here.
\end{remark}

\section{A constructive corollary}

If negative dependence narrows the fan, one can engineer it, and there is a clean continuous-marginal
construction that stays inside the $U_i\sim\mathrm{Unif}(0,1)$ model (unlike the discrete grid of
Theorem~\ref{thm:ends}). Take a one-per-stratum design: with $\pi$ a uniform random permutation of
$\{1,\dots,n\}$,
\begin{equation}
U_i=\frac{\pi(i)-1+V_i}{n}\quad(\text{Latin hypercube, }V_i\stackrel{\text{iid}}{\sim}\mathrm{Unif}(0,1)),
\qquad\text{or}\qquad U_i=\frac{\pi(i)-1+T}{n}\quad(\text{common jitter, }T\sim\mathrm{Unif}(0,1)).
\end{equation}
Each $U_i$ is exactly $\mathrm{Unif}(0,1)$ and the sample is exchangeable, with one point per stratum,
so the $k$-th order statistic lies in the $k$-th stratum and
$\mathbb E[U_{(k)}]=(k-\tfrac12)/n$, $\operatorname{Var}(U_{(k)})=1/(12n^2)$, a fan of order $1/n^2$
rather than $1/n$. (Note these continuous designs give mean $(k-\tfrac12)/n$, not the grid's exact
$k/(n+1)$.) More generally the classical variance-reduction designs apply: stratification never
increases variance \citep{cochran1977}, antithetic pairing helps for the monotone empirical CDF, and
determinantal or repulsive designs give negative association with faster-than-independent
concentration \citep{hough2006,bardenet2016}.

\begin{remark}[What is and is not admissible]
The marginal guarantee needs each calibration score exchangeable with the test score, not the
calibration scores independent of one another, so negative association is in principle compatible. But
the routes differ in what they cost. Score-dependent curation of the calibration set, the obvious way
to balance it, breaks exchangeability with the test point and voids the guarantee. A pooling that uses
all $n+1$ scores together (pooled ranks or de-meaning) preserves rank validity, but it is
transductive, full-conformal in flavour, not split conformal with a threshold fixed before the test
point is seen, and once the test score enters the pooled construction the realized-coverage variable
is no longer the fixed-threshold $c=U_{(k)}$ studied here. The safe message is the design one: a
pre-specified, exchangeability-preserving randomized design can lower the run-to-run variance of the
threshold without disturbing validity. What makes such a design admissible is not that it is merely
randomized and pre-specified, but that it preserves the exchangeability the validity argument needs;
covariate stratification qualifies only when it is external to the responses and applied symmetrically
across the $n+1$ points. Exact zero fan is the discrete rank/grid idealization; continuous-marginal
designs (Latin hypercube) give a small but nonzero fan, of order $1/n^2$.
\end{remark}

\begin{remark}[The copula viewpoint]\label{rmk:copula}
Because the scores are read on the uniform PIT scale, the joint law of $(U_1,\dots,U_n)$ is, by Sklar's
theorem, a copula $C$, with the marginal $F$ factored out entirely. So $\operatorname{Var}(c)$ is a
\emph{copula functional}: the fan is a copula invariant, identical for any score whose calibration
dependence is $C$. The fan coefficient is read off the bivariate copula diagonal,
$\bar c=\tfrac1{n(n-1)}\sum_{i\ne j}\big(C_{ij}(p,p)-p^2\big)$ at the operating quantile $p=1-\alpha$;
for the maximum ($k=n$) $\operatorname{Var}(c)$ is exactly a functional of the diagonal $u\mapsto
C(u,u)$, since $\mathbb P(U_{(n)}\le u)=C(u,u)$. The endpoints are the Fr\'echet bounds: independence
$\Pi$ gives the Beta fan, the comonotone copula $M$ ($U_1=\dots=U_n$) the maximal fan $\tfrac1{12}$;
the negative side has no dual, as the countermonotone lower bound is not a copula for $n>2$, which is
why the floor is only the most negative \emph{exchangeable} copula, $\rho=-1/(n-1)$.
Theorem~\ref{thm:pos} then says that the \emph{extendable} copulas, the exchangeable Archimedean
families (Clayton, Gumbel, Frank), each with a de Finetti frailty $W$ as directing variable, carry a
non-negative between-$Q$ term and widen the fan; the negative regime is the signed-de-Finetti corner
with no such mixture, which is why its contraction needed convex order rather than conditioning.
Finally, tail dependence acts through the copula's tails: a copula dependent in one tail pins the
order statistics there and fattens the opposite end, so lower-tail dependence (Clayton) inflates the
high, conformal quantile more than upper-tail dependence (Gumbel) --- the reverse of the naive guess,
as the companion demonstration shows.
\end{remark}

\section{Related work}

The $\mathrm{Beta}(k,n-k+1)$ law of calibration-conditional coverage and its variance are
\citet{vovk2012} and \citet{marques2024}; to our knowledge the papers cited here do not study the
dependence among calibration scores as a control variable for the realized-coverage variance. Negative association and its consequences are \citet{joagdev1983}; that negatively associated
sums are dominated in convex order by independent ones is \citet{shao2000}, and sampling without
replacement concentrates at least as tightly as with replacement \citep{serfling1974}. The total
positivity behind the single-crossing step is \citet{karlin1968}; that an order statistic of a
dependent sample need not be dispersively smaller than the independent one, which is why we claim
only the variance, is \citet{panja2022}; the connection between supermodular ordering and negative
dependence is \citet{christofides2004}; the Bahadur representation is \citet{bahadur1966}. Determinantal and repulsive designs are \citet{hough2006,bardenet2016} and
stratification is \citet{cochran1977}. The finite signed de Finetti representation is
\citet{kerns2006}, and the within-dataset, across-$x$ reading of its sign is the companion note
\citep{cotton_signed}; the information gap is \citep{cotton_marginally}. The contribution here is to
make the dependence the control variable for the \emph{variance} of realized coverage: the
leading-order fan coefficient, the exact law-of-total-variance decomposition on the positive side, the
exact aggregate contraction under negative association (Theorem~\ref{thm:agg}), and, for equicorrelated
normals, the exact and dispersively-monotone per-level law. The general single-level negative-association
contraction is left as a conjecture. The copula reading (Remark~\ref{rmk:copula}) rests on Sklar's
theorem \citep{sklar1959,nelsen2006}; copulas enter conformal prediction elsewhere through the
multivariate \emph{output} (e.g.\ \citealp{sunyu2024}), a different axis from the calibration-set
copula here.

\section*{Appendix: the quantile-slope reduction for Proposition~\ref{prop:dispmon}}

With $Z_{(k)}=aD+b\eta$, $a=\sqrt{1-\rho}$, $b=\sqrt{(1+(n-1)\rho)/n}$, the budget
$(n-1)a^2+nb^2=n$ holds and $da/d\rho=-1/(2a)$, $db/d\rho=(n-1)/(2nb)$. Under the natural coupling
(the same $D,\eta$ for all $\rho$), $\dot Y:=\partial_\rho Z_{(k)}=-D/(2a)+(n-1)\eta/(2nb)$, and the
quantile evolves by $\partial_\rho q_\rho(u)=\mathbb E[\dot Y\mid Y=q]=:M(q)$. Writing
$m(q)=\mathbb E[D\mid Y=q]$ and $b\,\mathbb E[\eta\mid Y=q]=q-a\,m(q)$,
\begin{equation}
M(q)=-\frac{m(q)}{2a}+\frac{n-1}{2nb^2}\big(q-a\,m(q)\big).
\end{equation}
Differentiating and using $\tfrac1{2a}+\tfrac{(n-1)a}{2nb^2}=\tfrac{nb^2+(n-1)a^2}{2anb^2}=\tfrac1{2ab^2}$
(the budget),
\begin{equation}
M'(q)=\frac{n-1}{2nb^2}-\frac{m'(q)}{2ab^2},
\end{equation}
so $M'(q)\ge0$, dispersive monotonicity ($\partial_\rho q_\rho$ increasing in $q$), is equivalent to
$m'(q)\le (n-1)a/n$. Tweedie's identity $m'(q)=a\operatorname{Var}(D\mid Y=q)/b^2$ turns this into the
posterior-variance bound $\operatorname{Var}(D\mid Y=q)\le (n-1)b^2/n$, which Lemma~\ref{lem:slc} and
the Brascamp--Lieb inequality deliver, as in Proposition~\ref{prop:dispmon}.

\medskip
\noindent\emph{The strong-log-concavity constant in Lemma~\ref{lem:slc}.} The constant $n/(n-1)$ is
tracked as follows. On the hyperplane $H=\{\sum_i w_i=0\}$ the centred vector $w$ has $-\log$ density
$\tfrac12\|w\|^2$ (the covariance $I-\tfrac1nJ$ restricts to the identity on $H$), so the measure is
$1$-strongly-log-concave there; restricting to the convex cone $\{w_1\le\dots\le w_n\}\cap H$ adds a
convex term to $-\log$ and keeps it $1$-strongly-log-concave. Strong log-concavity is preserved under
marginalization along a linear functional \citep{prekopa1973,brascamplieb1976,saumard2014}: for an
$m$-strongly-log-concave measure and $\ell(x)=\langle v,x\rangle$, the law of $\ell$ is
$(m/\|v\|^2)$-strongly-log-concave (the marginal along the unit direction $v/\|v\|$ is $m$-slc, and
rescaling the argument by $\|v\|$ scales the constant by $1/\|v\|^2$). Here $D=w_{(k)}=\langle e_k,w\rangle$
on the cone, with $v=\mathrm{proj}_H(e_k)=e_k-\tfrac1n\mathbf 1$ and $\|v\|^2=\tfrac{n-1}n$, so
$m=1$ gives the constant $n/(n-1)$ (the cone restriction only raises it). Numerically the realized
constant is larger still, $(\log g_D)''\le-2.5$ in the cases checked, so the bound holds with room.

\begin{thebibliography}{99}
\bibitem[Bahadur(1966)]{bahadur1966} Bahadur, R.~R. (1966). A note on quantiles in large samples.
\emph{The Annals of Mathematical Statistics} 37(3):577--580.
\bibitem[Bardenet and Hardy(2020)]{bardenet2016} Bardenet, R. and Hardy, A. (2020). Monte Carlo with
determinantal point processes. \emph{The Annals of Applied Probability} 30(1):368--417.
arXiv:1605.00361.
\bibitem[Christofides and Vaggelatou(2004)]{christofides2004} Christofides, T.~C. and Vaggelatou, E.
(2004). A connection between supermodular ordering and positive/negative association. \emph{Journal
of Multivariate Analysis} 88(1):138--151.
\bibitem[Cochran(1977)]{cochran1977} Cochran, W.~G. (1977). \emph{Sampling Techniques}, 3rd ed.
Wiley.
\bibitem[Cotton()]{cotton_marginally} Cotton, P. Marginally Useful: Formalizing the Information Gap
in Conformal Prediction. Companion paper.
\bibitem[Cotton()]{cotton_signed} Cotton, P. A Feynman--Wigner-Style Diagnostic for the Efficacy of
Conformal Prediction via Signed de Finetti Representations. Companion note.
\bibitem[de Finetti(1937)]{definetti1937} de Finetti, B. (1937). La pr\'evision: ses lois logiques,
ses sources subjectives. \emph{Annales de l'Institut Henri Poincar\'e} 7(1):1--68.
\bibitem[Hough et~al.(2006)]{hough2006} Hough, J.~B., Krishnapur, M., Peres, Y. and Vir\'ag, B.
(2006). Determinantal processes and independence. \emph{Probability Surveys} 3:206--229.
\bibitem[Joag-Dev and Proschan(1983)]{joagdev1983} Joag-Dev, K. and Proschan, F. (1983). Negative
association of random variables, with applications. \emph{The Annals of Statistics} 11(1):286--295.
\bibitem[Kerns and Sz\'ekely(2006)]{kerns2006} Kerns, G.~J. and Sz\'ekely, G.~J. (2006). De Finetti's
theorem for abstract finite exchangeable sequences. \emph{Journal of Theoretical Probability}
19(3):589--608.
\bibitem[Marques F.(2024)]{marques2024} Marques F., P.~C. (2024). Universal distribution of the
empirical coverage in split conformal prediction. \emph{Statistics \& Probability Letters} 219:110350
(2025). arXiv:2303.02770.
\bibitem[Serfling(1974)]{serfling1974} Serfling, R.~J. (1974). Probability inequalities for the sum
in sampling without replacement. \emph{The Annals of Statistics} 2(1):39--48.
\bibitem[Brascamp and Lieb(1976)]{brascamplieb1976} Brascamp, H.~J. and Lieb, E.~H. (1976). On
extensions of the Brunn--Minkowski and Pr\'ekopa--Leindler theorems, including inequalities for log
concave functions, and with an application to the diffusion equation. \emph{Journal of Functional
Analysis} 22(4):366--389.
\bibitem[Efron(2011)]{efron2011} Efron, B. (2011). Tweedie's formula and selection bias.
\emph{Journal of the American Statistical Association} 106(496):1602--1614.
\bibitem[Hansen and Tong(2026)]{hansen2026} Hansen, P.~R. and Tong, C. (2026). Tweedie's formula and
score-driven updating. Preprint, arXiv:2605.15902.
\bibitem[Karlin(1968)]{karlin1968} Karlin, S. (1968). \emph{Total Positivity, Vol.~1}. Stanford
University Press.
\bibitem[Lewis and Thompson(1981)]{lewis1981} Lewis, T. and Thompson, J.~W. (1981). Dispersive
distributions, and the connection between dispersivity and strong unimodality. \emph{Journal of
Applied Probability} 18(1):76--90.
\bibitem[Pr\'ekopa(1973)]{prekopa1973} Pr\'ekopa, A. (1973). On logarithmic concave measures and
functions. \emph{Acta Scientiarum Mathematicarum} 34:335--343.
\bibitem[Saumard and Wellner(2014)]{saumard2014} Saumard, A. and Wellner, J.~A. (2014).
Log-concavity and strong log-concavity: a review. \emph{Statistics Surveys} 8:45--114.
\bibitem[Panja et~al.(2022)]{panja2022} Panja, A., Kundu, P. and Pradhan, B. (2022). Dispersive and
star ordering of sample extremes from dependent random variables following the proportional odds
model. Preprint, arXiv:2006.04454.
\bibitem[Nelsen(2006)]{nelsen2006} Nelsen, R.~B. (2006). \emph{An Introduction to Copulas}, 2nd ed.
Springer.
\bibitem[Sklar(1959)]{sklar1959} Sklar, A. (1959). Fonctions de r\'epartition \`a $n$ dimensions et
leurs marges. \emph{Publications de l'Institut de Statistique de l'Universit\'e de Paris} 8:229--231.
\bibitem[Sun and Yu(2024)]{sunyu2024} Sun, S.~H. and Yu, R. (2024). Copula conformal prediction for
multi-step time series forecasting. \emph{ICLR}. arXiv:2212.03281.
\bibitem[Shao(2000)]{shao2000} Shao, Q.-M. (2000). A comparison theorem on moment inequalities
between negatively associated and independent random variables. \emph{Journal of Theoretical
Probability} 13(2):343--356.
\bibitem[Vovk(2012)]{vovk2012} Vovk, V. (2012). Conditional validity of inductive conformal
predictors. \emph{Asian Conference on Machine Learning}, PMLR 25:475--490.
\end{thebibliography}

\end{document}
